---
title: "hw5"
author: "HanSun, HanSun97"
date: "2020/4/10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
```

1. The complete dataset is too large. Download and import the alcohol sales data for Ames only:

```{r}
df <- readr::read_csv("2019_Iowa_Liquor_Sales.csv")
head(df)
```

2.Data cleaning:
a. extract geographic latitude and longitude from the variable Store Location

```{r}
Store_location <- separate(df, "Store Location", into = c("latitude", "longtitude"), sep = " (?=[^ ]+$)")
Store_location$latitude<- gsub("POINT \\(", "", Store_location$latitude)
Store_location$longtitude<- gsub("\\)", "", Store_location$longtitude)
```

b. make sure the variables are in the correct type, by converting incorrect variable types.

```{r}
Store_location$latitude <- as.numeric(Store_location$latitude)
Store_location$longtitude <- as.numeric(Store_location$longtitude)
Store_location$`Vendor Number` <- as.numeric(Store_location$`Vendor Number`)
Store_location$`Store Name` <- as.factor(Store_location$`Store Name`)
Store_location$`Category Name`<- as.factor(Store_location$`Category Name`)
```

c. extract year, month and day from the variable Date as separate columns.

```{r}
df_new <- Store_location %>%
  separate(Date, into = c("Month", "Day", "Year"))
head(df_new)
```

3. First overview:
a. Plot a scatterplot of lat and long of store locations.

```{r,fig.width=20, fig.height=10}
ggplot(df_new, aes(x=Store_location$longtitude, y=Store_location$latitude)) + geom_point()
```

b. Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown.

```{r,fig.width=10, fig.height=10}
df_new %>% 
  group_by(`Category Name`) %>% 
  ggplot(aes(x = `Category Name`, fill = `Volume Sold (Liters)`)) + geom_bar() + coord_flip()
```

4. Find the daily sales liquor sales in Ames in 2019. Summarize the number of sales, volume of liquor sold, and amount of money spent.

```{r}
df_2019 <- df_new %>%
  filter(Year == 2019) 
df_sales <- df_2019 %>%
  group_by(Day, Month) %>%
  summarize(Sales = sum(`Bottles Sold`), Vol = sum(`Volume Sold (Gallons)`), amount = sum(`Sale (Dollars)`))
df_sales
```

5. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Discover any trend with the help of geom_smooth. Describe any patterns.

```{r,fig.width=20, fig.height=5}
ggplot(df_2019, aes(x = Day, y = `Volume Sold (Gallons)`)) + geom_point() + facet_wrap(~Month)
```

From the graph, we get that the peak of each month is always between 10th to 20th. But the peak of April is 24th, the peak of May is 8th.

6. Find the dates for ISU football home games in Fall 2019. Can you see a pattern?

```{r,fig.width=20, fig.height=5}
df_sales %>% ggplot(aes(x = Day, y = Vol)) + geom_point() + facet_wrap(~Month)
```

The highest peak for the whole year is July 31th. The dates for ISU foot ball home games may in July 31th 2019.